Summary of CA Rice Breeding

### 2017-04-28 output/breed_paper.Rnw
Updates figures

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{Sweave}
\linespread{1.3}

\title{Summary of CA Rice Breeding}
\author{Matthew Espe}

@@

\maketitle

@@

\section{Background}

In general, rice yields in CA have been flat over the last 20+ years,
compared to the preceding period. Yet, annual reports from the CA
Cooperative Rice Research Foundation show that each released cultivar
has a yield advantage over the previous cultivars, typically on the
order of 4--6\%. This raises the question of how to reconcile these
facts. Namely, there are two possibilities:

\begin{enumerate}
\item{There is a mistake in the annual reports, and there is little or no
    yield advantage of newer cultivars over old.}
\item{The comparison to old cultivars is flawed, allowing an
    apparent yield increase to manifest without an accompanied
    increase in statewide yields.}
\end{enumerate}

\subsection{Motivating Evidence}

\begin{itemize}
\item At the year of release (1984), M-202 averaged 9,600 lbs acre\textsuperscript{-1}.
\item In 2015, the yield of M-209 was between 9,000-10,000 lbs acre\textsuperscript{-1}.
\item Therefore, there does not seem to have been an appreciable
  increase in yield over this period.
\end{itemize}

\subsection{Hypotheses}

\begin{enumerate}
  \item H1: After a cultivar has been released, its yield decreases over
    time.
  \item H2: Each cultivar release is recovering to the old yield benchmark, not
    increasing yield potential.
\end{enumerate}

\section{H1: Do yields decreases over time once a cultivar is released?}

\subsection{Methods}

@@

$$ yield = \alpha + \beta_{years\ in\ trial}* x_{years\ in\ trials} + \beta_{site} + \beta_{year} + \beta_{site:year} + \beta_{cultivar}$$

@@

\subsection{Results}

@@

<<fig=TRUE, echo=FALSE>>=
yrs_in <- -6:34
pp <- posterior_predict(yrs_fit, re.form = NA, newdata = data.frame(yrs_in_trial = yrs_in))
pp <-  pp * 1000 + 9000

@@

plot(yield_lb ~ yrs_in_trial, data = mod_data, pch = 16, col = rgb(0,0,0,0.25),
     xlab = "Year since entry",
     ylab = expression("Yield ( lbs acre"^-1~")"))
lines(x = yrs_in, y = ints[2,], lty = 2) 
polygon(x = c(yrs_in, rev(yrs_in)), y = c(ints[1,],rev(ints[3,])), col = rgb(0,0,0,0.2), border = NA)
text(20, 3500, bquote(.(round(fixef(yrs_fit)[2] * 1000, 1))~lbs~year^-1))

@ 
\caption{There is a small decrease in yield over the time since a
  cultivar was entered into the trial.}
\label{fig:yrs}
\end{figure}

As seen in Fig. \ref{fig:yrs}, there is a small decrease in yield
over time once a cultivar is released. This decrease is estimated to be
\Sexpr{round(fixef(yrs_fit)[2] * 1000, 1)} lbs
year\textsuperscript{-1}. The oldest cultivar in the trials has lost
an estimated \Sexpr{round(fixef(yrs_fit)[2] * 1000 * 36, 1)} lbs
acre\textsuperscript{-1} since release.

\section{H2: Do newer cultivars have a yield advantage?}

\subsection{Methods}

$$ yield = \alpha + \beta_{release\ year}*x_{release\ year} +
\beta_{site} + \beta_{year} + \beta_{site:year} + \beta_{cultivar}$$

\subsection{Results}

@@

<<fig=TRUE, echo=FALSE>>=
yrs <- 1980:2015
pp <- posterior_predict(release_fit, re.form = NA, newdata = data.frame(release_yr_c = yrs - 2000))
pp <-  pp * 1000 + 9000

plot(yield_lb ~ release_yr, data = mod_data, pch = 16, col = rgb(0,0,0,0.25),
     xlab = "Year of Release",
     ylab =  expression("Yield ( lbs acre"^-1~")"))
lines(x = yrs, y = ints[2,], lty = 2) 
polygon(x = c(yrs, rev(yrs)), y = c(ints[1,],rev(ints[3,])), col = rgb(0,0,0,0.2), border = NA)
text(1990, 3500, bquote(.(round(fixef(release_fit)[2] * 1000, 1))~lbs~year^-1))
@

\caption{There is an estimated increase in yield over time, but it is
  smaller than the previously estimated 54 lbs
  year\textsuperscript{-1}}
\label{fig:release}
\end{figure}

@@

@@

\section{Complications}

@@

\begin{figure}
<<fig=TRUE, fig.width = 8, fig.height = 5, out.width = '80%', echo=FALSE>>=
tars <- c("M202","M206","M209")
par(mfrow = c(1,3),
    mar = c(4,0,2,0),
    oma = c(4,4,0.5,0.5))

invisible(sapply(tars, function(x){
    plot(yield_lb ~ year, data = mod_data[(mod_data$id == x & mod_data$site == "RES"),],
         xlim = c(1995, 2016), ylim = c(6000,12000),
         pch = 16, main = x,
         ylab = "",
         yaxt = "n",
         xlab = "Year")
    if(x == "M202"){
        axis(2)
        mtext(expression("Yield ( lbs acre"^-1~")"), 2, line = 2.25, cex.lab = 0.6)
    }
    abline(lm(yield_lb ~ year, data = mod_data[(mod_data$id == x & mod_data$site == "RES"),]), lty = 2)

}))

@
\caption{The performance of three major cultivars at the RES site over
time.}
\label{fig:res_yrs}
\end{figure}

<<fig=TRUE, echo=FALSE>>=
par(mfrow = c(1,3),
    mar = c(4,0,2,0),
    oma = c(4,4,0.5,0.5))

invisible(sapply(tars, function(x){
    plot(yield_lb ~ yrs_in_trial,
         data = mod_data[(mod_data$id == x & mod_data$site == "RES"),],
         xlim = c(-6, 36), ylim = c(6000,12000),
         pch = 16, main = x,
         ylab = "",
         yaxt = "n",
         xlab = "Years after release")
    if(x == "M202"){
        axis(2)
        mtext(expression("Yield ( lbs acre"^-1~")"), 2, line = 2.25, cex.lab = 0.6)
    }
    abline(lm(yield_lb ~ yrs_in_trial,
              data = mod_data[(mod_data$id == x & mod_data$site == "RES"),]), lty = 2)
}))

@
\caption{The performance of three major cultivars at the RES site
  since the year each cultivar was released.}
\label{fig:res_since_release}
\end{figure}

>>>>

\documentclass[12pt,]{elsarticle}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile, textcomp}
\usepackage{setspace}
\usepackage{etoolbox}
\doublespacing
\usepackage{lineno}
\usepackage{lscape} %% Harvard
\bibliographystyle{model2-names}\biboptions{authoryear}
\pagenumbering{gobble}
%% For figure numbers in supplementary info
%% from http://bytesizebio.net/2013/03/11/adding-supplementary-tables-and-figures-in-latex/
\newcommand{\beginsupplement}{%
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}%
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}%
}
\usepackage{Sweave}
\SweaveOpts{prefix.string=graphics/report}

@@

\begin{frontmatter}

\title{Rice yield improvement by plant breeding are offset by
  decreases over time}
\author[ucd]{Matthew B. Espe\corref{corresauthor}}
\cortext[corresauthor]{Corresponding author} \ead{mespe@ucdavis.edu}

\address[ucd]{Dept. of Plant Sciences, University of California --
Davis, Davis CA 95616}

\begin{abstract}

  Using the variety trial data, we find consistent estimates of the
  reduction of yield once a variety is released, estimated to be between
  -12 to -22 kg per year. This effect persists using all data, only
  named/released cultivars, or only released cultivars only at the Rice
  experiment station. This effect is offset by yield improvements over
  the years, estimated to be small but positive effect of

\end{abstract}

\begin{keyword} rice, yield potential, temperature
\end{keyword}

\end{frontmatter}

\textbf{Abbreviations}

CA: California

CI: Credible Interval

%\linenumbers

@@

\section{Introduction}\label{intro}

\setlength{\parindent}{1cm}

Constraints on ariable land are increasing simultaneous with the need
to increase total production to meet growing demand, which has
necessitated harvesting more grain per unit land area. Historically,
agricultural research has been successful in staving off the Mathesian
disaster through continued yield improvement of staple crops, however
more and more production systems are experiencing plateaus in grain
yield \citep{Peng2010}. If continued improvements per unit land area
are no longer possible, increasing the area under cultivation will be
needed to meet demand, which carries with it undesirable ecological
implications. Therefore, it is of critical importance to better
understand why grain yields have not continued to increase.

CA rice production represents one such rice production system. Rice
yields in CA during the period from 1920 -- 1990 increased rapidly, but
since 1990 have increased only slowly if at all (Fig. 1
\ref{fig:avg_yields}). However, continued improvement of grain yield
as the result of plant breeding has been demonstrated since 1990
\citep{rrb}, with each new cultivar having a yield advantage over the
previous cultivars. Yet, these same reports show stagnant yields; for
example, at the year of release (1984), cultivar 'M-202' averaged
9,600 lbs acre\textsuperscript{-1}. In 2015, the yield of 'M-209' was
between 9,000 -- 10,000 lbs acre\textsuperscript{-1}. We sought to better
understand how grain yields have increased in the breeders' fields
without a corresponding increase in yield in the farmers' fields. Two
competing hypotheses were tested: (1) Grain yields have not improved
since 1990, or (2) grain yield improvements have been offset by an
associated decrease in yield once a cultivar was selected.

\begin{figure}[h]
<<fig=TRUE, echo=FALSE>>=
CA_data = read.csv("../data/CA_rice_yield.csv", stringsAsFactors = FALSE)
CA_data$Value = as.numeric(gsub(",", "", CA_data$Value))
plot(Value * 1.12 ~ Year, CA_data,
     type = "l",
     xlab = "Year",
     ylab = expression("Average grain yield (kg ha"^-1~")"))

@
\caption{Average rice yields over the period \Sexpr{min(CA_data$Year)}
  -- \Sexpr{max(CA_data$Year)} for California, USA. Rapid gains in
    grain yield were experienced from approximately 1950 -- 1990, but
    annual gains since 1990 have been slower.}
\label{fig:state_yield}
\end{figure}

In the course of investigating the drivers of yield in this system, we
were compelled to explore how yields could increase with each release
of improved cultivars, yet average yield for the region remain
flat. Specifically, we offered two competing hypotheses to explain
this phenomenon: (1) the yield advantage offered by newly developed
cultivars is not greater than previous cultivars, or (2) yield
increases gained through plant breeding are being offset by decreases
in yield of released cultivars.

Yet, annual reports from the CA Cooperative Rice Research Foundation
show that each released cultivar has a yield advantage over the
previous cultivars, typically on the order of 4 -- 6\%. This raises the
question of how to reconcile these facts. Namely, there are two
possibilities:

\section{Methods}\label{methods}

\subsection{Site description}

\subsection{Statistical analysis}

@@

\begin{equation}
 yield = \alpha + \beta_{years\ in\ trial}* x_{years\ in\ trials} + \beta_{site} + \beta_{year} + \beta_{site:year} + \beta_{cultivar}
\end{equation}

@@

\begin{equation}
yield = \alpha + \beta_{release\ year}*x_{release\ year} +
\beta_{site} + \beta_{year} + \beta_{site:year} + \beta_{cultivar}
\end{equation}

\section{Results}\label{results}

\subsection{Yield improvements over time}

@@

<<fig=TRUE, echo=FALSE>>=
yrs <- 1980:2015
pp <- posterior_predict(release_fit, re.form = NA, newdata = data.frame(release_yr_c = yrs - 2000))
pp <-  (pp * 1000 + 9000) * 1.12

@@

res = mod_data$site == "RES"
plot_data = as.data.frame(tapply(mod_data$yield_lb[res] * 1.12,
                                 mod_data$id[res], mean))
plot_data$se = tapply(mod_data$yield_lb[res] * 1.12,
                      mod_data$id[res],
                      function(x)
                          sd(x)/sqrt(length(x)))

plot_data$tmp = tapply(mod_data$release_yr[res],
                       mod_data$id[res], unique)
colnames(plot_data) = c("yield_kg", "yield_se", "release_yr")

plot(yield_kg ~ release_yr, data = plot_data, pch = 16, #col = rgb(0,0,0,0.25),
     xlab = "Release year",
     ylab = expression("Grain yield ( kg ha"^-1~")"),
     ylim = c(8000, 13000),
     xlim = c(1980,2019))
abline(a = (((fixef(RES_rel_fit)[1] * 1000) + 9000) * 1.12) -
           (2000 * fixef(RES_rel_fit)[2] * 1000),
       b = fixef(RES_rel_fit)[2] * 1000, col = "grey")
with(plot_data, segments(x0 = release_yr, y0=yield_kg + yield_se,
                         y1 = yield_kg - yield_se))
text(x = plot_data$release_yr, y = plot_data$yield_kg,
     labels = gsub("M", "M-", row.names(plot_data)),
     pos = 4, cex = 0.8)

# ints <- apply(pp, 2, quantile, p = c(0.05, 0.5, 0.95))
# polygon(x = c(yrs_in, rev(yrs_in)), y = c(ints[1,],rev(ints[3,])), col = rgb(0,0,0,0.2), border = NA)
# text(0, 7000, bquote(.(round(fixef(RES_yrs_fit)[2] * 1000, 1))~lbs~year^-1))

# plot(yield_lb ~ release_yr, data = mod_data, pch = 16, col = rgb(0,0,0,0.25),
#      xlab = "Year of Release",
#      ylab =  expression("Yield ( lbs acre"^-1~")"))
# lines(x = yrs, y = ints[2,], lty = 2) 
# polygon(x = c(yrs, rev(yrs)), y = c(ints[1,],rev(ints[3,])), col = rgb(0,0,0,0.2), border = NA)
text(1990, 8000, bquote(.(round(fixef(release_fit)[2] * 1000 * 1.12, 1))~kg~ha^-1~year^-1))
@

\caption{The yield increase over time of medium grain rice varieties
  in California, USA. There is an estimated
  increase in yield over time, but it is smaller than a previously
  estimate of 54 lbs year\textsuperscript{-1}. This estimate is
  potentially misleading, because it only
  considers the average yield of each variety during the period 1995
  -- 2015, and the older varieties have experiencing greater losses
  since being released.}
\label{fig:release}
\end{figure}

\subsection{Yield decreases over time}

% \begin{figure}[h]

% <<fig=TRUE, echo=FALSE>>=
% yrs_in <- -6:34
% pp <- posterior_predict(yrs_fit, re.form = NA, newdata = data.frame(yrs_in_trial = yrs_in))
% pp <-  (pp * 1000 + 9000) * 1.12

% ints <- apply(pp, 2, quantile, p = c(0.05, 0.5, 0.95))

% plot_data = as.data.frame(tapply(mod_data$yield_lb * 1.12, mod_data$yrs_in_trial, mean))
% plot_data$se = tapply(mod_data$yield_lb * 1.12, mod_data$yrs_in_trial, function(x)
%     sd(x)/sqrt(length(x)))
% plot_data$tmp = as.numeric(row.names(plot_data))
% colnames(plot_data) = c("yield_kg", "yield_se", "yrs_in_trial")

% plot(yield_kg ~ yrs_in_trial, data = plot_data, pch = 16, #col = rgb(0,0,0,0.25),
%      xlab = "Year since entry",
%      ylab = expression("Yield ( kg ha"^-1~")"),
%      ylim = c(8700, 11200))
% with(plot_data, segments(x0 = yrs_in_trial, y0=yield_kg + yield_se,
%                          y1 = yield_kg - yield_se))
% abline(a = ((fixef(yrs_fit)[1] * 1000) + 9000) * 1.12, b = fixef(yrs_fit)[2] * 1000)
% # lines(x = yrs_in, y = ints[2,], lty = 2) 
% # polygon(x = c(yrs_in, rev(yrs_in)), y = c(ints[1,],rev(ints[3,])), col = rgb(0,0,0,0.2), border = NA)
% text(30, 4500, bquote(.(round(fixef(yrs_fit)[2] * 1000, 1))~lbs~year^-1))

% @ 
% \caption{There is a small decrease in yield over the time since a
  % cultivar was entered into the trial using the full data
  % set. However, the effect is small (\Sexpr{round(fixef(yrs_fit)[2] * 1000, 1)}
  % lbs year\textsuperscript{-1}) in relation to the variability
  % between sites and years.}
% \label{fig:yrs}
% \end{figure}

<<echo=FALSE>>=
all_CI <- round(posterior_interval(yrs_fit, pars = "yrs_in_trial") * 1000, 1)
RES_CI <- round(posterior_interval(RES_yrs_fit, pars = "yrs_in_trial") * 1000, 1)
@

As seen in Fig. \ref{fig:yrs}, there is a small decrease in yield
over time once a cultivar is released, though the estimated effect is
small compared to the variability in the data set. This decrease is estimated to be
\Sexpr{round(fixef(yrs_fit)[2] * 1000, 1)} lbs
year\textsuperscript{-1}. The oldest cultivar in the trials has lost
an estimated \Sexpr{round(fixef(yrs_fit)[2] * 1000 * 36 * 1.12, 1)} kg ha\textsuperscript{-1} since release. Using just data from the Rice
Experiment Station (RES), the estimated effect is more drastic
(an estimated \Sexpr{round(fixef(RES_yrs_fit)[2] * 1000 * 1.12, 1)} kg
ha\textsuperscript{-1} year\textsuperscript{-1}), but with a wider credible interval
(\Sexpr{RES_CI[1]} -- \Sexpr{RES_CI[2]} vs \Sexpr{all_CI[1]} --
\Sexpr{all_CI[2]}).

@@

<<fig=TRUE, echo=FALSE>>=
yrs_in <- -7:35
pp <- posterior_predict(RES_yrs_fit, re.form = NA, newdata = data.frame(yrs_in_trial = yrs_in))
pp <-  (pp * 1000 + 9000) * 1.12

RES = res = mod_data$site == "RES"
plot_data = as.data.frame(tapply(mod_data$yield_lb[res] * 1.12,
                                 mod_data$yrs_in_trial[res], mean))
plot_data$se = tapply(mod_data$yield_lb[res] * 1.12,
                      mod_data$yrs_in_trial[res],
                      function(x)
                          sd(x)/sqrt(length(x)))

plot_data$tmp = as.numeric(row.names(plot_data))
colnames(plot_data) = c("yield_kg", "yield_se", "yrs_in_trial")

plot(yield_kg ~ yrs_in_trial, data = plot_data, pch = 16, #col = rgb(0,0,0,0.25),
     xlab = "Year since entry",
     ylab = expression("Grain yield ( kg ha"^-1~")"),
     ylim = c(7000, 13000),
     xlim = c(-5,33))
with(plot_data, segments(x0 = yrs_in_trial, y0=yield_kg + yield_se,
                         y1 = yield_kg - yield_se))
abline(a = ((fixef(RES_yrs_fit)[1] * 1000) + 9000) * 1.12,
       b = fixef(RES_yrs_fit)[2] * 1000)
ints <- apply(pp, 2, quantile, p = c(0.05, 0.5, 0.95))
#polygon(x = c(yrs_in, rev(yrs_in)), y = c(ints[1,],rev(ints[3,])), col = rgb(0,0,0,0.2), border = NA)
text(0, 7000, bquote(.(round(fixef(RES_yrs_fit)[2] * 1000 * 1.12, 1))~kg~ha^-1~year^-1))

@
\caption{The relationship between yield and the years since entry for
  medium grain rice varieties planted at the Rice Experiment Station
  (RES), California, USA. RES is considered
  the best managed site, and has the most plots per year of any of the
  sites. The estimated decrease over time is estimated to be stronger
  using only these data (\Sexpr{round(fixef(RES_yrs_fit)[2] * 1000, 1)}
  lbs year\textsuperscript{-1}), primarly due to lower variability.}
\label{fig:RES_yrs}
\end{figure}

\begin{figure}
<<fig=TRUE, fig.width = 10, fig.height = 5, out.width = '80%', echo=FALSE>>=
tars <- c("M202","M206")
par(mfrow = c(1, length(tars)),
    mar = c(4,0,2,0),
    oma = c(4,4,0.5,0.5))

invisible(sapply(tars, function(x){
    plot(yield_lb * 1.12 ~ year,
         data = mod_data[(mod_data$id == x & mod_data$site == "RES"),],
         xlim = c(1995, 2016), ylim = c(5000,14000),
         pch = 16, main = gsub("M", "M-", x),
         ylab = "",
         yaxt = "n",
         xlab = "Year")
    if(x == "M202"){
        axis(2)
        mtext(expression("Yield ( kg ha"^-1~")"), 2, line = 2.25, cex.lab = 0.6)
    }
    # abline(lm(yield_lb ~ year, data = mod_data[(mod_data$id == x & mod_data$site == "RES"),]), lty = 2)
    abline(h = mean(mod_data$yield_lb[(mod_data$id == x & mod_data$site == "RES")] * 1.12),
           lty = 2)
}))

@
\caption{The performance of two major cultivars at the CA Rice
  Experiment Station (RES) over the period 1995 -- 2015. The dashed line is the average
  yield for each cultivar over the period which it was planted at the
  RES. The newer cultivar, M-209, appears to have a slight yield
  advantage over the older cultivar, M-202.}
\label{fig:res_yrs}
\end{figure}

<<fig=TRUE, echo=FALSE>>=
par(mfrow = c(1, length(tars)),
    mar = c(4,0,2,0),
    oma = c(4,4,0.5,0.5))

invisible(sapply(tars, function(x){
    plot(yield_lb*1.12 ~ yrs_in_trial,
         data = mod_data[(mod_data$id == x & mod_data$site == "RES"),],
         xlim = c(-6, 36), ylim = c(5000,14000),
         pch = 16, main = gsub("M","M-", x),
         ylab = "",
         yaxt = "n",
         xlab = "Years after release")
    if(x == "M202"){
        axis(2)
        mtext(expression("Yield ( kg ha"^-1~")"), 2, line = 2.25, cex.lab = 0.6)
    }
    abline(lm(yield_lb*1.12 ~ yrs_in_trial,
              data = mod_data[(mod_data$id == x & mod_data$site == "RES"),]), lty = 2)
}))

@
\caption{The performance of two major cultivars at the CA Rice
  Experiment Station (RES) since the year each cultivar was
  released. Both M-202 and M-206 have similar intercepts, suggesting
  similar yield at the time of release and an apparent yield advantage
  of the newer cultivar, M-209, might be explained by a yield loss of
  M-202 over time.}
\label{fig:res_since_release}
\end{figure}

@@

% \begin{figure}[h]

% <<fig=TRUE, echo=FALSE>>=
% yrs <- 1980:2015
% pp <- posterior_predict(RES_rel_fit, re.form = NA,
%                         newdata = data.frame(release_yr_c = yrs - 2000))
% pp <-  pp * 1000 + 9000

% ints <- apply(pp, 2, quantile, p = c(0.05, 0.5, 0.95))

% plot(yield_lb ~ release_yr, data = mod_data[RES,], pch = 16, col = rgb(0,0,0,0.25),
%      xlab = "Year of Release",
%      ylab =  expression("Yield ( lbs acre"^-1~")"))
% lines(x = yrs, y = ints[2,], lty = 2) 
% polygon(x = c(yrs, rev(yrs)), y = c(ints[1,],rev(ints[3,])), col = rgb(0,0,0,0.2), border = NA)
% text(2010, 4500, bquote(.(round(fixef(RES_rel_fit)[2] * 1000, 1))~lbs~year^-1))
% @

% \caption{The effect of year of release on yield for only data from the
%   Rice Experiment Station (RES). There is an estimated
%   increase in yield over time, but it is smaller than the previously estimated 54 lbs
%   year\textsuperscript{-1}}
% \label{fig:RES_release}
% \end{figure}

\section{Discussion}

\subsection{Complications}

@@

@@

\section{References}\label{references}

\setlength{\parindent}{-0.2in} \setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}

\bibliography{breeding}

\section{Supplemental tables and
figures}\label{supplemental-tables-and-figures}

\beginsupplement

### 2017-04-28 output/summary.Rnw
Updates figures

@@

In general, rice yields in CA have been flat over the last 20+ years,
compared to the preceding period. Yet, annual reports from the CA
Cooperative Rice Research Foundation show that each released cultivar
has a yield advantage over the previous cultivars, typically on the
order of 4--6\%. This raises the question of how to reconcile these
facts. Namely, there are two possibilities:

@@

plot(yield_lb ~ yrs_in_trial, data = mod_data, pch = 16, col = rgb(0,0,0,0.25),
     xlab = "Year since entry",
     ylab = expression("Yield ( lbs acre"^-1~")"))
lines(x = yrs_in, y = ints[2,], lty = 2) 
polygon(x = c(yrs_in, rev(yrs_in)), y = c(ints[1,],rev(ints[3,])), col = rgb(0,0,0,0.2), border = NA)
text(20, 3500, bquote(.(round(fixef(yrs_fit)[2] * 1000, 1))~lbs~year^-1))

@ 
\caption{There is a small decrease in yield over the time since a
  cultivar was entered into the trial.}
\label{fig:yrs}
\end{figure}

As seen in Fig. \ref{fig:yrs}, there is a small decrease in yield
over time once a cultivar is released. This decrease is estimated to be
\Sexpr{round(fixef(yrs_fit)[2] * 1000, 1)} lbs
year\textsuperscript{-1}. The oldest cultivar in the trials has lost
an estimated \Sexpr{round(fixef(yrs_fit)[2] * 1000 * 36, 1)} lbs
acre\textsuperscript{-1} since release.

>>>>

\section{Summary}

Using the variety trial data, we find consistent estimates of the
reduction of yield once a variety is released, estimated to be between
-12 to -22 kg per year. This effect persists using all data, only
named/released cultivars, or only released cultivars only at the Rice
experiment station. This effect is offset by yield improvements over
the years, estimated to be small but positive effect of

@@

Increased constraints on ariable land simultaneous with the need to
increase total production to meet growing demand have necessitated
harvesting more grain per unit land area over time. Historically,
agricultural research has been successful in staving off the Mathesian
disaster through continued yield improvement, however more and more
production systems are experiencing
plateaus in grain yield \citep{cassman}. If continued improvements per
unit land area are no longer possible, increasing the total land under
cultivation will be needed to meet demand, which carries with it
undesirable ecological implications. Therefore, it is of critical
importance to better understand why grain yields have not continued to
increase.

In general, rice yields in CA have been flat over the last 20+ years,
compared to the preceding period, despite a demonstrated improvement
in grain yield via plant breeding \citep{rrb}. In the course of
investigating the drivers of yield in this system, we were compelled
to explore how yields could increase with each release of improved
cultivars, yet average yield for the region remain flat. Specifically,
we offered two competing hypotheses to explain this phenomenon: (1)
the yield advantage offered by newly developed cultivars is not
greater than previous cultivars, or (2) yield increases gained through
plant breeding are being offset by decreases in yield of released
cultivars.

Yet, annual reports from the CA
Cooperative Rice Research Foundation show that each released cultivar
has a yield advantage over the previous cultivars, typically on the
order of 4--6\%. This raises the question of how to reconcile these
facts. Namely, there are two possibilities:

@@

plot(yield_lb ~ yrs_in_trial, data = mod_data, pch = 16, col = rgb(0,0,0,0.25),
     xlab = "Year since entry",
     ylab = expression("Yield ( lbs acre"^-1~")"))
lines(x = yrs_in, y = ints[2,], lty = 2) 
polygon(x = c(yrs_in, rev(yrs_in)), y = c(ints[1,],rev(ints[3,])), col = rgb(0,0,0,0.2), border = NA)
text(30, 4500, bquote(.(round(fixef(yrs_fit)[2] * 1000, 1))~lbs~year^-1))

@ 
\caption{There is a small decrease in yield over the time since a
  cultivar was entered into the trial using the full data
  set. However, the effect is small (\Sexpr{round(fixef(yrs_fit)[2] * 1000, 1)}
  lbs year\textsuperscript{-1}) in relation to the variability
  between sites and years.}
\label{fig:yrs}
\end{figure}

<<echo=FALSE>>=
all_CI <- round(posterior_interval(yrs_fit, pars = "yrs_in_trial") * 1000, 1)
RES_CI <- round(posterior_interval(RES_yrs_fit, pars = "yrs_in_trial") * 1000, 1)
@

As seen in Fig. \ref{fig:yrs}, there is a small decrease in yield
over time once a cultivar is released, though the estimated effect is
small compared to the variability in the data set. This decrease is estimated to be
\Sexpr{round(fixef(yrs_fit)[2] * 1000, 1)} lbs
year\textsuperscript{-1}. The oldest cultivar in the trials has lost
an estimated \Sexpr{round(fixef(yrs_fit)[2] * 1000 * 36, 1)} lbs
acre\textsuperscript{-1} since release. Using just data from the Rice
Experiment Station (RES), the estimated effect is more drastic
(an estimated \Sexpr{round(fixef(RES_yrs_fit)[2] * 1000, 1)} lbs
year\textsuperscript{-1}), but with a wider credible interval
(\Sexpr{RES_CI[1]} -- \Sexpr{RES_CI[2]} vs \Sexpr{all_CI[1]} --
\Sexpr{all_CI[2]}).

<<fig=TRUE, echo=FALSE>>=
yrs_in <- -6:34
pp <- posterior_predict(RES_yrs_fit, re.form = NA, newdata = data.frame(yrs_in_trial = yrs_in))
pp <-  pp * 1000 + 9000

ints <- apply(pp, 2, quantile, p = c(0.05, 0.5, 0.95))
RES <- mod_data[,'site'] == "RES"

plot(yield_lb ~ yrs_in_trial, data = mod_data[RES,], pch = 16, col = rgb(0,0,0,0.25),
     xlab = "Year since entry",
     ylab = expression("Yield ( lbs acre"^-1~")"))
lines(x = yrs_in, y = ints[2,], lty = 2) 
polygon(x = c(yrs_in, rev(yrs_in)), y = c(ints[1,],rev(ints[3,])), col = rgb(0,0,0,0.2), border = NA)
text(30, 5000, bquote(.(round(fixef(RES_yrs_fit)[2] * 1000, 1))~lbs~year^-1))

@
\caption{The relationship between yield and years since entry for only
a single site, the Rice Experiment Station (RES). RES is considered
the best managed site, and has the most plots per year of any of the
sites. The estimated decrease over time is estimated to be stronger
using only these data (\Sexpr{round(fixef(RES_yrs_fit)[2] * 1000, 1)}
lbs year\textsuperscript{-1}), primarly due to lower variability.}
\label{fig:RES_yrs}
\end{figure}

\begin{figure}
<<fig=TRUE, fig.width = 8, fig.height = 5, out.width = '80%', echo=FALSE>>=
tars <- c("M202","M206")
par(mfrow = c(1, length(tars)),
    mar = c(4,0,2,0),
    oma = c(4,4,0.5,0.5))

invisible(sapply(tars, function(x){
    plot(yield_lb ~ year, data = mod_data[(mod_data$id == x & mod_data$site == "RES"),],
         xlim = c(1995, 2016), ylim = c(5000,14000),
         pch = 16, main = x,
         ylab = "",
         yaxt = "n",
         xlab = "Year")
    if(x == "M202"){
        axis(2)
        mtext(expression("Yield ( lbs acre"^-1~")"), 2, line = 2.25, cex.lab = 0.6)
    }
    abline(lm(yield_lb ~ year, data = mod_data[(mod_data$id == x & mod_data$site == "RES"),]), lty = 2)

}))

@
\caption{The performance of two major cultivars at the RES site over
time. There is evidence that a cultivar's performance decreases over
time.}
\label{fig:res_yrs}
\end{figure}

\begin{figure}

<<fig=TRUE, echo=FALSE>>=
par(mfrow = c(1, length(tars)),
    mar = c(4,0,2,0),
    oma = c(4,4,0.5,0.5))

invisible(sapply(tars, function(x){
    plot(yield_lb ~ yrs_in_trial,
         data = mod_data[(mod_data$id == x & mod_data$site == "RES"),],
         xlim = c(-6, 36), ylim = c(5000,14000),
         pch = 16, main = x,
         ylab = "",
         yaxt = "n",
         xlab = "Years after release")
    if(x == "M202"){
        axis(2)
        mtext(expression("Yield ( lbs acre"^-1~")"), 2, line = 2.25, cex.lab = 0.6)
    }
    abline(lm(yield_lb ~ yrs_in_trial,
              data = mod_data[(mod_data$id == x & mod_data$site == "RES"),]), lty = 2)
}))

@
\caption{The performance of two major cultivars at the RES site
  since the year each cultivar was released. Both M-202 and M-206 have
  similar intercepts, suggesting similar yield at the time of release.}
\label{fig:res_since_release}
\end{figure}
